---
title: " Covid-19 Vaccine Tweets and Side Effects Report"
author: "Group L"
output:
  html_document:
    toc: true
    self_contained: true
    keep_md: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message=FALSE))
```


# Overview  

For this project, we explored two datasets -- tweets about the Covid-19 vaccine and the reported side effects of the vaccine.   
Firstly, we looked into what people are discussing when they tweet about the Covid-19 vaccine and explored their feelings about it. This would give us some insight into people's general attitudes towards it.    
Secondly, we focused on the adverse reactions reported from 2020-12-01 to 2021-3-31. The visualization aims to provide an insight into who are the people reporting side effects and how do they compare to the general population; the most common reported symptoms etc.    
Altogether, our project aims to provide insights to guide more appropriate actions in promoting Covid-19 vaccination, as well as more effective actions in informing and relieving side effects.


# Data

- All COVID-19 Vaccines Tweets
https://www.kaggle.com/gpreda/all-covid19-vaccines-tweets
- Vaccine Adverse Event Reporting System (VAERS)1 established by the Food and Drug Administration (FDA) and the Centers for Disease Control and Prevention (CDC) 
https://vaers.hhs.gov/data/datasets.html?
- Allocations of Covid-19 vaccines produced by Pfizer and Moderna into different States by CDC https://data.cdc.gov/Vaccinations/COVID-19-Vaccine-Distribution-Allocations- by-Juris/saz5-9hgg https://data.cdc.gov/Vaccinations/COVID-19-Vaccine-Distribution-Allocations- by-Juris/b7pe-5nws
- COVID-19 Vaccination Demographic Data. Vaccination by age. https://www.cdc.gov/coronavirus/2019-ncov/vaccines/distributing/demographics-vaccination-data.html
- State-by-state data on COVID-19 vaccinations in the United States 
https://ourworldindata.org/us-states-vaccinations

<br/>
<br/>

# PART Ⅰ    
<br/>     

# Tweets about Covid-19 Vaccine      
<br/>    

## What Do People Tweet about Covid-19 Vaccines ? 

```{r include=FALSE}
library(tidyverse)
library(tm)
library(stopwords)
library(tidytext)
library(wordcloud)
library(tidyverse)
library(hunspell)
library(data.table)
library(readxl)
library(igraph)
library(ggthemes)
library(ggplot2)
library(igraph)
library(tibble)
library(plotly)
library(network)
library(DT)
library(wordcloud)
library(sna)
library(visNetwork)
library(ggrepel)
library(networkD3)
library(gridExtra)
# library(RColorBrewer)

tweet <- fread("./data/vaccination_all_tweets_ansi.csv")
```


```{r}
load("./data/term_stemmed_all.RData")
```

```{r eval=FALSE, include=FALSE}
# str(tt)
# all oringinal tweets

# clean text
tt <- tweet[,c("id","text")]
tt <- rename(tt, "doc_id" = 'id')
tt$text <- as.character(tt$text)
# make a corpus
library(tm)
tt_source <- DataframeSource(tt)
tt_corpus <- VCorpus(tt_source)
tt_corpus


library(stopwords)
# preprocessing function----
clean_corpus <- function(corpus){
  corpus <- tm_map(corpus, removePunctuation)
  corpus <- tm_map(corpus, removeWords, c(stopwords(source = "smart")))
  corpus <- tm_map(corpus, removeNumbers)
  corpus <- tm_map(corpus, stripWhitespace)
  return(corpus)
}
# preprocessing
tt_clean <- clean_corpus(tt_corpus)
tt_dtm <- DocumentTermMatrix(tt_clean)

library(tidytext)
tt_tidy <- tidy(tt_dtm)
# mannually remove residual unwanted words
tt_tidy <- tt_tidy %>% 
  filter(grepl("[a-z]*\\…",term) == F) %>% 
  filter(grepl("http([a-z]*[A-Z]*[0-9]*)*",term) == F) %>% 
  # filter(grepl("([A-Z]*[a-z]*)*'([A-Z]*[a-z]*)*'",term) == F) %>% # 去除
  filter(term != "the") %>% 
  filter(term != "this") %>% 
  group_by(term) %>% 
  mutate(n=n()) %>% 
  arrange(desc(n))

tt_tidy$term2 <- hunspell::hunspell_stem(tt_tidy$term,"en_US")
ttt <- tt_tidy
ttt$term2 <- lapply(ttt$term2, function(x) if(identical(x, character(0))) NA_character_ else x)
stem1 <- ttt %>% 
  ungroup() %>% 
  mutate(index = row_number())%>% 
  unnest(term2)
stem1 <- stem1 %>% mutate(term3 = ifelse(is.na(term2),term,term2))
stemmed_all <- stem1[c(1,7,3)] %>% 
  rename(term = term3) %>% 
  group_by(term) %>% 
  mutate(n = sum(count)) %>% 
  arrange(desc(n))
save(stemmed_all,file = "term_stemmed_all.RData")
```

```{r echo=FALSE,dpi= 300,out.width = '70%',out.height = '70%'}
# plot word cloud
library(wordcloud)
set.seed(377)
library(RColorBrewer)
# reds <- brewer.pal(5, "RdPu")
purple_orange <- brewer.pal(10, "RdYlBu")
stemmed_all_p <- unique(stemmed_all
                    [c(2,4)]) # dataset for plotting
# save(stemmed_p,file = "term_stemmed.RData")
wordcloud(stemmed_all_p$term,stemmed_all_p$n, scale=c(3,1),min.freq = 600,
          max.words = 200, color=purple_orange)
```

This word cloud includes the popular keywords (appeared more than 600 times) used in tweeting about Covid-19 vaccine.    
The most commonly mentioned word is of course "vaccine", followed by "moderna" and "covid", while "pfizer" and "pfizerbiontech" are much smaller.   
We can also see some common keywords seemingly describing experiences "dose","receive","today", suggesting many of these tweets may be recording people's vaccination experiences.   
There's some discussion about China (“china” and “chinese”)， because Chinese also manufacture and hand out Covid-19 vaccine.
It's worth noting that "sore" and "side" also appears a lot, so maybe a couple of people suffering side effects.


## A Co-occurrence Network of Keywords

```{r include=FALSE}
library(tidyverse)
library(tidyr)
library(rlist)
library(igraph)
library(ggnetwork)
library(networkD3)
```

```{r eval=FALSE, include=FALSE,dpi= 300,out.width = '70%',out.height = '70%'}
# make an co-occurence matrix
list1k <- stemmed_all[,c(2,4)] %>% unique() %>% arrange(desc(n)) %>% head(1000) #n>=67
top1k <- stemmed_all %>% filter(term %in% list1k$term)
words <- top1k[,c(2,1)]

cooc <- merge(words,words,by='document')
cooc <- cooc %>%
  filter(term.x != term.y) %>% # remove one's relation with it self
  count(term.x,term.y) # times of co-ocurrence
cooc <- cooc %>% arrange(desc(n))
# remove replecate records (undirected edges)
cooc <- cooc %>% rename(from=term.x) %>% rename(to=term.y)
cooc2 <- cooc %>% filter(n>=100)
library(tidyverse)
dedup <- cooc2 %>%
  mutate(normalized = map2_chr(from, to, ~paste(sort(c(.x, .y)),
                                                collapse = ""))) %>%
  group_by(normalized,n) %>%
  summarise(from = first(from),
            to = first(to)) %>%
  ungroup() %>%
  select(-normalized) %>%
  arrange(desc(n))
uniedge <- dedup[,c(2,3,1)]
# save(cooc, file = "coocedge.Rdata")
# save(uniedge, file = "uniquedges.Rdata")
# library(readr)
# write_csv(cooc, path = "coocedge.csv")
```

```{r include=FALSE}
load("./data/coocedge.Rdata")
load("./data/uniquedges.Rdata")
```


```{r include=FALSE}
# make a graph----
edges <- as.matrix(uniedge)
g <- graph_from_data_frame(edges, directed = F, vertices = NULL)
# # g
# # calculate indegree and outdegree (centrality)
# V(g)$deg <- degree(g, mode = "total", loops = F, normalized = FALSE)
# # add weight(n) as edge width
# E(g)$width <- E(g)$n
# # as_data_frame(g, what="edges") 
# # g
# # plot(g)
```

```{r include=FALSE}
# # prepare data for plotting----
# library(ggnetwork)
# set.seed(12)
# # class(g)
# # summary(g)
# df <- ggnetwork(g,layout = with_kk())
# # df
# att <- rename(unique(stemmed_all[,c(2,4)]),name = term)
# df <- left_join(df,att,by="name")
# df$width <- as.numeric(df$width)/100
# 
#find clusters
wc <- cluster_walktrap(g,weights = E(g)$width)  # find "communities"
# # class(c)
# # sizes(wc)
members <- membership(wc)
mm <- as.data.frame(cbind(names(members),members)) %>% rename(.,name=V1) %>% rename(.,cluster=members)
# # ignore tiny groups
# mm_b <- mm %>% group_by(cluster) %>% 
#   mutate(n = n()) %>% 
#   mutate(group = ifelse(n>=5,n,1))
# # summary(as.factor(mm_b$group))
# # name the groups
# mm_b$group <- as.factor(mm_b$group) # numeric to factor
# mm_b$group <- ordered(mm_b$group, levels = c(1,7,23,37,212), labels = c("undefined","group1","group2","group3","group4"))
# mm_b$group <- as.character(mm_b$group)
# df_m <- merge(mm_b,df,by='name')
# 
# # # plot the clusters
# # cp <- unique(df_m[,c(1,4)])
# # library(ggthemes)
# # cp %>% ggplot(aes(name,group)) +
# #   geom_jitter(height = 0.2) +
# #   theme_clean() +
# #   labs(title = "Automatically Detected Clusters",
# #        tag = "Figure 2",
# #        y = "cluster") +
# #   theme(axis.text.x = element_blank(),axis.title.x = element_blank()) +
# #   scale_colour_manual(values = c("blue", "green", "red")) +
# #   labs(color='Party')
# 
# ```
# 
# ```{r eval=FALSE, include=FALSE}
# # plotting network
# ggplot(df_m, aes(x, y, xend = xend, yend = yend)) +
#   geom_edges(aes(size = width),alpha = 0.2,color="seashell3",curvature = 0.2) +
#   geom_nodes(data=subset(df_m, group=="undefined"),aes(size=n.y), color = "gray3", alpha = 0.5) + # terms not in big clusters
#   geom_nodes(data=subset(df_m, group=="group1"),aes(size=n.y), color = "darkred", alpha = 0.5) + # group1
#   geom_nodes(data=subset(df_m, group=="group2"),aes(size=n.y), color = "blue4", alpha = 0.5) + # group2
#   geom_nodes(data=subset(df_m, group=="group3"),aes(size=n.y), color = "purple", alpha = 0.5) + # group3
#   geom_nodes(data=subset(df_m, group=="group4"),aes(size=n.y), color = "orange", alpha = 0.5) + # group4
#   geom_nodetext_repel(aes(label = name),size = 2.5) + 
#   theme_blank()
# # no text for central nodes ---- try interactive graph

```

```{r echo=FALSE}
library(networkD3)
# Convert to object suitable for networkD3
g_d3 <- igraph_to_networkD3(g, group = members)
# Create force directed network plot
forceNetwork(Links = g_d3$links, Nodes = g_d3$nodes, 
             Source = 'source', Target = 'target', 
             NodeID = 'name', Group = 'group',
             fontSize = 50, fontFamily = "Times New Roman",
             zoom = T)
# no specifying link width (would look messy)
```

There are four major clusters detected – one is the major group with two central points – “vaccine” and “covid”; another is one around “moderna” the manufacturer, which may come from tweets reporting new progresses of moderna vaccine; another one is more dispersed with three centers – “today”, russia” and “antario”, which may come from those focus on vaccine exportation news；the other one at the intercept is more disperse and doesn’t have a central term.    
The clusters are interwoven together, but can offer some hins on different popular topics. Readers can freely explore the network and look for the relevant words they are interested in.


```{r eval=FALSE, include=FALSE}
## Sentiment Analysis

twi <- read_excel("./data/DV_sentiment_NEW.xlsx")
twi <- as.data.table(twi)

twi<-twi%>%
mutate(Sentiment=ifelse(sentiment>= 0.05,"positive",ifelse(sentiment <= -0.05,"negative","neutral")))

twi<-twi%>%
mutate(twi_date=as.Date(twi$date,format="%Y-%m-%d"))%>%
mutate(td= format(as.Date(twi_date, "%m/%d/%Y"),"%Y-%m"))
save(twi,file="twi.RData")
```


```{r include=FALSE}
load("./data/twi.RData")
```

## Sentiment State Distribution of Tweets
```{r include=FALSE}
three <- twi%>%
  select(Sentiment)%>%
  group_by(Sentiment)%>%
  count(Sentiment)
```

```{r echo=FALSE,dpi= 300,out.width = '70%',out.height = '70%'}
ggplot(three, aes(x=Sentiment, y=n, fill=Sentiment)) +
  geom_bar(stat="identity")+theme_minimal()+labs(x = "Sentiment Categories", y = "Number of Tweets",
       title = "Sentiment State Distribution of Tweets")+theme(legend.position="none")+ 
scale_fill_manual("legend", values = c("positive" = "#FF6666", "neutral" = "#faead3", "negative" = "#106d9c"))
```

We clean the text data of all the tweets about vaccines, and then we apply vader sentiment analysis, so we classify all tweets into three categories: positive tweets, neutral tweets and negative tweets.

Now we can see people's attitude towards vaccines.

We can find from the bar chart that most tweets about vaccines are neutral one or positive one. Negative sentiment is not widely available.


```{r echo=F,include=FALSE}
num<-twi%>%
  select(twi_date,Sentiment)%>%
  group_by(twi_date,Sentiment)%>%
  count(twi_date,Sentiment)
```


```{r echo=FALSE,,dpi= 300,out.width = '70%',out.height = '70%'}
ggplot(num, aes(x = twi_date, y = n,colour= Sentiment)) +scale_x_date(breaks = "25 day")+ theme_minimal()+geom_line(size=1.2)+labs(x = "Date of Tweets", y = "Number of Tweets",
       title = "Timeline showing Sentiment of Tweets about COVID-19 Vaccines") + scale_color_manual(values=c("#106d9c", "yellow",  "#FF6666"))
```

Trends over time of numbers of tweets posted of three sentiment types are similar, maybe because there is no special events affects people's attitude.


```{r include=FALSE}
pos <-  read_csv("./data/pos.csv")
neu <-  read_csv("./data/neu.csv")
neg <-  read_csv("./data/neg.csv")
```


```{r echo=F,dpi= 300,out.width = '70%',out.height = '70%'}
set.seed(12)
layout(matrix(c(1, 2), nrow=2), heights=c(1, 4))
par(mar=rep(0, 4))
plot.new()
text(x=0.5, y=0.5, "Positive")
wordcloud(words = pos$tokenize, freq = pos$n, min.freq = 1,
          scale=c(3,1), max.words = 200, color=purple_orange)
```

```{r echo=F,dpi= 300,out.width = '70%',out.height = '70%'}
set.seed(12)
layout(matrix(c(1, 2), nrow=2), heights=c(1, 4))
par(mar=rep(0, 4))
plot.new()
text(x=0.5, y=0.5, "Negative")
wordcloud(words = neg$tokenize, freq = neg$n, min.freq = 1,
          scale=c(3,1), max.words = 200, color=purple_orange)
```

```{r echo=FALSE,dpi= 300,out.width = '70%',out.height = '70%'}
set.seed(12)
layout(matrix(c(1, 2), nrow=2), heights=c(1, 4))
par(mar=rep(0, 4))
plot.new()
text(x=0.5, y=0.5, "Neutral")
wordcloud(words = neu$tokenize, freq = neu$n, min.freq = 1,
          scale=c(3,1), max.words = 200, color=purple_orange)
```

We select top 50 common words for each types of tweets.

After seeing the common words of positive, neutral and negative  tweets, we find people share their happiness about the arrival of vaccines and give positive feedback after receiving a shot in positive tweets; neutral tweets are just  objective statements of vaccines news or information; people worry about the side effect of vaccines and whether vaccines will work in negative tweets.


## Portrait of Popular Tweets and Users
```{r include=FALSE}
favor<-twi%>%
  select(favorites,Sentiment)%>%
  arrange(-favorites)%>%  
  slice(1:15)
```


```{r echo=FALSE,dpi= 300,out.width = '70%',out.height = '70%'}
ggplot(data=favor,aes(y=as.factor(favorites), x=favorites,fill=Sentiment) ) + 
  geom_bar(stat="identity")+labs(x = "Number of Times it was Favorited", y = "Tweets",
       title = "Sentiment State of Top 15 Popular Tweets base on favorited times ")+theme_minimal()+  theme(
        axis.text.y=element_blank(),axis.ticks.y=element_blank())+ geom_text(aes(label=paste0(favorites)), vjust=0.5,hjust=1)+ 
scale_fill_manual("legend", values = c("positive" = "#FF6666", "neutral" = "#faead3", "negative" = "#106d9c"))
```

We can see sentiment attribute of popular tweets(base on favorited times). Most of top 15 popular tweets are neutral one or positive one, which means that people didn't show a preference for negative tweets.

```{r include=FALSE}
retwe<-twi%>%
  select(retweets,Sentiment)%>%
  arrange(-retweets)%>%  
  slice(1:15)
```


```{r echo=FALSE,dpi= 300,out.width = '70%',out.height = '70%'}
ggplot(data=retwe,aes(y=as.factor(retweets), x=retweets,fill=Sentiment) ) + 
  geom_bar(stat="identity")+labs(x = "Number of Times it was Retweeted", y = "Tweets",
       title = "Sentiment State of Top 15 Popular Tweets base on retweeted times ")+theme_minimal()+ theme(
        axis.text.y=element_blank(),axis.ticks.y=element_blank())+ geom_text(aes(label=paste0(retweets)), vjust=0.5,hjust=1)+ 
scale_fill_manual("legend", values = c("positive" = "#FF6666", "neutral" = "#faead3", "negative" = "#106d9c"))
```

We can see sentiment attribute of popular tweets(base on retweeted times). Most of top 15 popular tweets still are neutral one or positive one, which means that people didn't show a preference for retweeting negative tweets and maybe kept positive attitude towards effect of vaccines.

```{r include=FALSE}
follow <-  read_csv("./data/follow.csv")%>%
  mutate(Sentiment=sent_ca)

```


```{r echo=FALSE,dpi= 300,out.width = '70%',out.height = '70%'}
ggplot(follow, aes(x=factor(user_name), fill=Sentiment))+
  geom_bar(width=0.7)+
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = 70, hjust = 0.5, 
                              vjust = 0.5,color = "black",size=9))+labs(x = "User Name", y = "Number of Tweets",
       title = "Sentiment State of Tweets of Top 10 Users who have most followers ")+ 
scale_fill_manual("legend", values = c("positive" = "#FF6666", "neutral" = "#faead3", "negative" = "#106d9c"))
```

We also check tweets of popular users(base on number of followers they have), because they have great influence among the public. Most of these users mainly post objective statements of vaccines news or information and they show more positive sentiment than negative sentiment.





```{r}
packages <- c("devtools","knitr","tidyverse","widgetframe","readr",
              "wordcloud", "base64enc", "tidytext", 
              "RWeka","stats","manifestoR","readtext",
              "rvest", "stringr", 
              "SnowballC", "plotrix", "tidyr", "tidytext", "stats", 
              "dendextend", "ggthemes",
              "httr","jsonlite", "DT", "textdata", "ggmap","maptools","mapproj","rgeos","rgdal",
              "RColorBrewer", "stringr","scales", "leaflet", 'leafpop', "ggthemes", "ggtext", "wordcloud")

packages <- lapply(packages, FUN = function(x) {
  if(!require(x, character.only = TRUE)) {
    install.packages(x)
  library(x, character.only = TRUE)
  }
}
)
```
